---
title: "Hierarchical Agglomerative Single Linkage Clustering Simulation"
author: "MA Mingyu 14110562D"
date: "4/5/2017"
output: pdf_document
---
#COMP4433 Assignment 2 Question 3d
##derek.ma@connect.polyu.hk [derek.ma](http://derek.ma)

#Set Up
Import data, delete first column, set initial cluster centers to first two records and set k is equal to 2.
```{r setup, include=TRUE}
library(dplyr)
#PREPARE DATA IN USE
data <- read.csv("~/Google Drive/_DM/2_Assignments/Ass2/data_q3.csv", stringsAsFactors = FALSE)
#delete first column for this specific case becasue it is not a data record
data <- data[,2:length(colnames(data))]
```

# Preprocessing and Normalization
```{r preprocessing}
#Normalization
min1 <- min(data$B)
max1 <- max(data$B)
min2 <- min(data$C)
max2 <- max(data$C)
min3 <- min(data$D)
max3 <- max(data$D)
min4 <- min(data$E)
max4 <- max(data$E)
min5 <- min(data$F)
max5 <- max(data$F)
min6 <- min(data$G)
max6 <- max(data$G)

data_normalized <- data %>%
  mutate(BN = (B - min1)/(max1-min1)) %>%
  mutate(CN = (C - min2)/(max2-min2)) %>%
  mutate(DN = (D - min3)/(max3-min3)) %>%
  mutate(EN = (E - min4)/(max4-min4)) %>%
  mutate(FN = (F - min5)/(max5-min5)) %>%
  mutate(GN = (G - min6)/(max6-min6))

data_normalized <- data_normalized[,c(7,8,9,10,11,12)]
colnames(data_normalized) <- c("B","C","D","E","F","G")
data <- data_normalized
```

# Set Parameters
In this program, a clusters data frame will be used to save which nodes are in which clusters. Initial clusters distribution is saved in `clusters` data frame. After n iterations, the clusters distribution information will be saved in `clustersn` data frame.
```{r setParas}
# Create initial proximity matrix
proximityMatrix <- data.frame(matrix(NA, ncol = nrow(data), nrow = nrow(data)))
colnames(proximityMatrix) <- (1:nrow(data))
# Create a data frame to save all clusters
clusters <- data.frame(matrix(NA, ncol = nrow(data), nrow = nrow(data)))
colnames(clusters) <- (1:nrow(data))
colnames(clusters) <- paste("cluster", colnames(clusters), sep = "_")
clusters[1,] <- 1:nrow(data)
```

#Basic Function
## Distance Function
```{r distance,include}
distance <- function(vector1, vector2){
  #Euclidean distance
  #Input: two vectors of data with same length
  #Input example: c(1,2,3); c(2,3,4)
  count <- 0
  for (i in 1:length(vector1)){
    count = count + (vector1[i] - vector2[i])^2
  }
  count^(1/2)
}
```

## indexOfMin Function - Find row and column number for the minimum proximity in the proximity matrix
```{r function2}
indexOfMin <- function(objectsData){
  #set all 0 values to NA
  valueMinInRows <- rep(NA,nrow(objectsData))
  colMinInRows <- rep(NA,nrow(objectsData))
  objectsData[objectsData == 0] <- Inf
  objectsData[is.na(objectsData)] <- Inf
  for (i in 1:nrow(objectsData)){
    colMinInRows[i] <- which.min(as.numeric(objectsData[i,]))
    valueMinInRows[i] <- min(objectsData[i,])
  }
  row <- which.min(valueMinInRows)
  col <- colMinInRows[row]
  c(row,col)
}
```

## minOfMatrix Function - return the minimum value except 0 and NA in the matrix
```{r function3}
minOfMatrix <- function(objectsData){
  #set all 0 values to NA
  valueMinInRows <- rep(NA,nrow(objectsData))
  colMinInRows <- rep(NA,nrow(objectsData))
  objectsData[objectsData == 0] <- Inf
  objectsData[is.na(objectsData)] <- Inf
  for (i in 1:nrow(objectsData)){
    colMinInRows[i] <- which.min(as.numeric(objectsData[i,]))
    valueMinInRows[i] <- min(objectsData[i,])
  }
  min(valueMinInRows)
}
```

## updateClusters Function - Based on the minimum proximity information, combine clusters
```{r function-updateClusters}
updateClusters <- function(index1,index2,clustersData){
  newClustersData <- clustersData
  baseIndex <- sum(!is.na(clustersData[,index2]))
  addIndex <- sum(!is.na(clustersData[,index1]))
  for (i in 1:addIndex){
    newClustersData[baseIndex+i,index2] <- clustersData[i,index1]
  }
  newClustersData[,-index1]
}
```

## updateProximity Function - calculate proximity between each pair of clusters and generate proximity matrix
```{r initialProximityMatrix}
updateProximity <- function(clustersData, objectsData, proximityData){
  newProximityData <- data.frame(matrix(NA,nrow=ncol(clustersData),ncol=ncol(clustersData)))
  colnames(newProximityData) <- colnames(clustersData)
  for (irow in 1:ncol(clustersData)){
    for (icol in irow:ncol(clustersData)){
      # For each pair of clusters
      currentMinDist <- 0
      for (pointClu1 in 1:sum(!is.na(clustersData[,irow]))){
        for (pointClu2 in 1:sum(!is.na(clustersData[,icol]))){
          # For each pair of points
          dist_value <- distance(
            as.numeric(objectsData[clustersData[pointClu1,irow],]),
            as.numeric(objectsData[clustersData[pointClu2,icol],])
          )
          if (dist_value < currentMinDist || currentMinDist == 0){
            currentMinDist <- dist_value
          }
        }
      }
      # save min distance between these two clusters
      newProximityData[irow,icol] <- currentMinDist
    }
  }
  newProximityData
}
```

# Iteration
For each iteration, following steps will be implemented:
1 - Calculate proximity between each pair of clusters and update proximityMatrix
2 - Find out where is the minimum proximity in the proximity matrix
3 - Combine clusters and update the clusters table
For each iteration, following tabls will be printed to answer the questions:
d(1) - updated distance matrix: `proximityMatrixn`(n is the times of iterations)
d(2) - partitions produced: `clustersn` can show the partitions for each iteration
d(2) - final dendrogram: attached after this program

## Iteration Round 1
```{r merge1}
proximityMatrix1 <- updateProximity(clusters,data,proximityMatrix)
combineIndex <- indexOfMin(proximityMatrix1)
minProximity <- minOfMatrix(proximityMatrix1)
clusters1 <- updateClusters(combineIndex[1],combineIndex[2],clusters)
proximityMatrix1
minProximity
clusters1
```

## Iteration Round 2
```{r merge2}
proximityMatrix2 <- updateProximity(clusters1,data,proximityMatrix1)
combineIndex <- indexOfMin(proximityMatrix2)
minProximity <- minOfMatrix(proximityMatrix2)
clusters2 <- updateClusters(combineIndex[1],combineIndex[2],clusters1)
proximityMatrix2
minProximity
clusters2
```

## Iteration Round 3
```{r merge3}
proximityMatrix3 <- updateProximity(clusters2,data,proximityMatrix2)
combineIndex <- indexOfMin(proximityMatrix3)
minProximity <- minOfMatrix(proximityMatrix3)
clusters3 <- updateClusters(combineIndex[1],combineIndex[2],clusters2)
proximityMatrix3
minProximity
clusters3
```

## Iteration Round 4
```{r merge4}
proximityMatrix4 <- updateProximity(clusters3,data,proximityMatrix3)
combineIndex <- indexOfMin(proximityMatrix4)
minProximity <- minOfMatrix(proximityMatrix4)
clusters4 <- updateClusters(combineIndex[1],combineIndex[2],clusters3)
proximityMatrix4
minProximity
clusters4
```

## Iteration Round 5
```{r merge5}
proximityMatrix5 <- updateProximity(clusters4,data,proximityMatrix4)
combineIndex <- indexOfMin(proximityMatrix5)
minProximity <- minOfMatrix(proximityMatrix5)
clusters5 <- updateClusters(combineIndex[1],combineIndex[2],clusters4)
proximityMatrix5
minProximity
clusters5
```

## Iteration Round 6
```{r merge6}
proximityMatrix6 <- updateProximity(clusters5,data,proximityMatrix5)
combineIndex <- indexOfMin(proximityMatrix6)
minProximity <- minOfMatrix(proximityMatrix6)
clusters6 <- updateClusters(combineIndex[1],combineIndex[2],clusters5)
proximityMatrix6
minProximity
clusters6
```

## Iteration Round 7
```{r merge7}
proximityMatrix7 <- updateProximity(clusters6,data,proximityMatrix6)
combineIndex <- indexOfMin(proximityMatrix7)
minProximity <- minOfMatrix(proximityMatrix7)
clusters7 <- updateClusters(combineIndex[1],combineIndex[2],clusters6)
proximityMatrix7
minProximity
clusters7
```

## Iteration Round 8
```{r merge8}
proximityMatrix8 <- updateProximity(clusters7,data,proximityMatrix7)
combineIndex <- indexOfMin(proximityMatrix8)
minProximity <- minOfMatrix(proximityMatrix8)
clusters8 <- updateClusters(combineIndex[1],combineIndex[2],clusters7)
proximityMatrix8
minProximity
clusters8
```

## Iteration Round 9
```{r merge9}
proximityMatrix9 <- updateProximity(clusters8,data,proximityMatrix8)
combineIndex <- indexOfMin(proximityMatrix9)
minProximity <- minOfMatrix(proximityMatrix9)
clusters9 <- updateClusters(combineIndex[1],combineIndex[2],clusters8)
proximityMatrix9
minProximity
clusters9
```

## Iteration Round 10
```{r merge10}
proximityMatrix10 <- updateProximity(clusters9,data,proximityMatrix9)
combineIndex <- indexOfMin(proximityMatrix10)
minProximity <- minOfMatrix(proximityMatrix10)
clusters10 <- updateClusters(combineIndex[1],combineIndex[2],clusters9)
proximityMatrix10
minProximity
clusters10
```

## Iteration Round 11
```{r merge11}
proximityMatrix11 <- updateProximity(clusters10,data,proximityMatrix10)
combineIndex <- indexOfMin(proximityMatrix11)
minProximity <- minOfMatrix(proximityMatrix11)
clusters11 <- updateClusters(combineIndex[1],combineIndex[2],clusters10)
proximityMatrix11
minProximity
clusters11
```

## Iteration Round 12
```{r merge12}
proximityMatrix12 <- updateProximity(clusters11,data,proximityMatrix11)
combineIndex <- indexOfMin(proximityMatrix12)
minProximity <- minOfMatrix(proximityMatrix12)
clusters12 <- updateClusters(combineIndex[1],combineIndex[2],clusters11)
proximityMatrix12
minProximity
clusters12
```

## Iteration Round 13
```{r merge13}
proximityMatrix13 <- updateProximity(clusters12,data,proximityMatrix12)
combineIndex <- indexOfMin(proximityMatrix13)
minProximity <- minOfMatrix(proximityMatrix13)
clusters13 <- updateClusters(combineIndex[1],combineIndex[2],clusters12)
proximityMatrix13
minProximity
clusters13
```

## Iteration Round 14
```{r merge14}
proximityMatrix14 <- updateProximity(clusters13,data,proximityMatrix13)
combineIndex <- indexOfMin(proximityMatrix14)
minProximity <- minOfMatrix(proximityMatrix14)
clusters14 <- updateClusters(combineIndex[1],combineIndex[2],clusters13)
proximityMatrix14
minProximity
clusters14
```
Now all nodes are in one cluster. The final dengrogram has been formed.

```{r verify,include=FALSE}
plot(hclust(dist(data),method="single",members = NULL))
```
