---
title: "Information_Gain_Computation"
author: "MA Mingyu 14110562D"
date: "4/4/2017"
output: pdf_document
---
#COMP4433 Assignment 2
##derek.ma@connect.polyu.hk [derek.ma](http://derek.ma)
This R markdown file can be found at https://github.com/derekmma/data-mining-algorithms/blob/master/apriori_simulation.Rmd

#Initialization of Training Data
```{r setup, include=TRUE}
#init demo data to a data frame
training_set <- data.frame(
  sex=c("F","F","M","M","F","F","M","F","M","F","F","F","M","M","F","F","M","F","M","F"),
  age=c("Y","M","S","M","Y","S","Y","M","Y","M","M","M","Y","S","S","Y","S","Y","M","S"),
  married=c("Y","N","N","Y","Y","Y","N","Y","N","Y","Y","Y","Y","N","Y","N","N","Y","N","Y"),
  income=c("H","H","M","M","M","L","H","L","M","M","H","H","L","M","M","L","H","H","M","L"),
  plan=c("A","C","B","B","C","B","C","C","A","C","C","A","B","A","A","C","C","B","B","A"),
  renew=c("Y","Y","N","Y","Y","N","N","Y","N","Y","Y","Y","N","N","N","Y","N","Y","Y","N")
  )
```

#Functions to Calculate Entropy and Informtion Gain
```{r entropy}
entropy <- function(dataset,targetFeature){
  #input: the target feature column after selected
  #out: the entropy under this condition
  target <- dataset[,targetFeature]
  allValues <- unique(target)
  n <- length(target)
  entropyValue <- 0
  for (value in allValues){
    p <- (length(subset(target, target == value))/n)
    entropyValue = entropyValue - p*log2(p)
  }
  entropyValue
}
```

```{r info-gain}
infoGain <- function(dataset,feature,targetFeature){
  infoGain_value <- entropy(dataset,targetFeature)
  currentColumn <- dataset[,feature]
  allValues <- unique(currentColumn)
  n <- length(currentColumn)
  for (value in allValues){
    p <- (length(subset(currentColumn, currentColumn == value))/n)
    entropyTemp <- entropy(subset(dataset, dataset[,feature] == value), targetFeature)
    infoGain_value <- infoGain_value - p*entropyTemp
  }
  infoGain_value
}
```

# Best Split Selection - First Round

In ID3, information gain is used to choose the splitting attribute.
Calculate information gain for each attribute to select the root split.

```{r first-split-selection}
infoGain_1 <- data.frame(
  sex = infoGain(training_set,"sex","renew"),
  age = infoGain(training_set,"age","renew"),
  married = infoGain(training_set,"married","renew"),
  income = infoGain(training_set,"income","renew"),
  plan = infoGain(training_set,"plan","renew")
  )
infoGain_1
```

According to the calculation result for infomation gain, initial split is on age, because it has the highest information gain.

# Split - First Round
```{r first-split}
training_set_ageY <- subset(training_set, training_set$age == "Y")
training_set_ageM <- subset(training_set, training_set$age == "M")
training_set_ageS <- subset(training_set, training_set$age == "S")
training_set_ageY
training_set_ageM
training_set_ageS
```

## Check Stopping Criteria
We found that for records of age "Middle", all of the customers will renew and for records of age "Senior", all of the customer will not renew. These two branches match the stopping criteria. We do not need to split these two branches any more. They are pure already.

# Best Split Selection - Second Round

Now we are going to find the best split feature for the branch with age "Young".
```{r second-split-selection}
infoGain_2 <- data.frame(
  sex = infoGain(training_set_ageY,"sex","renew"),
  married = infoGain(training_set_ageY,"married","renew"),
  income = infoGain(training_set_ageY,"income","renew"),
  plan = infoGain(training_set_ageY,"plan","renew")
  )
infoGain_2
```

We can find that "sex" feature has highest information gain. Then "sex" should be selected as next level split node.

# Split - Second Round
```{r second-split}
training_set_ageY_sexM <- subset(training_set_ageY, training_set_ageY$sex == "M")
training_set_ageY_sexF <- subset(training_set_ageY, training_set_ageY$sex == "F")
training_set_ageY_sexM
training_set_ageY_sexF
```

## Check Stopping Criteria
All samples for the given nodes belong to the same class. Thus the split action terminates and the decision has been formed.

```{r verification, include=FALSE}
library(tree)
tree <- tree(renew~sex+age+married+income+plan, data=training_set)
plot(tree)
text(tree, pretty=0)
```