---
title: "kmeans"
author: "MA Mingyu 14110562D"
date: "4/4/2017"
output:
  pdf_document: default
  html_document: default
---
#COMP4433 Assignment 2 Question 3a
##derek.ma@connect.polyu.hk [derek.ma](http://derek.ma)

#Set Up
Import data, delete first column, set initial cluster centers to first two records and set k is equal to 2.
```{r setup, include=TRUE}
library(dplyr)
#PREPARE DATA IN USE
data_original <- read.csv("~/Google Drive/_DM/2_Assignments/Ass2/data_q3.csv", stringsAsFactors = FALSE)
#delete first column for this specific case becasue it is not a data record
data_original <- data_original[,2:length(colnames(data_original))]
#SET VALUE OF K
k <- 2
data <- data_original
#SET INITIAL CLUSTER CENTERS
centers <- data[0,]
#colnames(centers) <- colnames(data)
#for (i in 1:k){
  #set first two record as initial centers
  #centers[i,] <- data[i,]
#}
```

```{r preprocessing}
#Normalization
min1 <- min(data$B)
max1 <- max(data$B)
min2 <- min(data$C)
max2 <- max(data$C)
min3 <- min(data$D)
max3 <- max(data$D)
min4 <- min(data$E)
max4 <- max(data$E)
min5 <- min(data$F)
max5 <- max(data$F)
min6 <- min(data$G)
max6 <- max(data$G)

data_normalized <- data %>%
  mutate(BN = (B - min1)/(max1-min1)) %>%
  mutate(CN = (C - min2)/(max2-min2)) %>%
  mutate(DN = (D - min3)/(max3-min3)) %>%
  mutate(EN = (E - min4)/(max4-min4)) %>%
  mutate(FN = (F - min5)/(max5-min5)) %>%
  mutate(GN = (G - min6)/(max6-min6))

data_normalized<-data_normalized[,c(7,8,9,10,11,12)]
colnames(data_normalized) <- c("B","C","D","E","F","G")
data <- data_normalized
centers <- data[c(1,2),]
```

# Basic Function 1 - Distance Function
In this case, Euclidean Distance is used to calculate the dissimilarities.
```{r distance}
distance <- function(vector1, vector2){
  #Euclidean distance
  #Input: two vectors of data with same length
  #Input example: c(1,2,3); c(2,3,4)
  count <- 0
  for (i in 1:length(vector1)){
    count = count + (vector1[i] - vector2[i])^2
  }
  count^(1/2)
}
```

# Basic Function 2 - Compare Similarity and Assign Objects to Clusters
In this function, each record can be decivded belong to which clusters.
```{r assignObject}
assign <- function(objectsData, centersData){
  #OUTPUT a data frame with new cluster information
  #FOR EACH RECORDS
  result <- objectsData %>% 
    mutate(cluster = NA)
  for (i in 1:nrow(objectsData)){
    whichCenter <- 0
    currentMinDist <- -1
    #COUNT DISSIMILARITY BETWEEN IT AND CENTERS
    for (j in 1:nrow(centersData)){
      distValue <- distance(
        as.numeric(objectsData[i,]),
        as.numeric(centersData[j,]))
      if (distValue < currentMinDist || whichCenter == 0){
        #FOUND CENTER WITH SMALLER DISSIMILARITY
        whichCenter <- j
        currentMinDist <- distValue
      }
    }
    #SET THIS CENTER AS CLUSTER
    result[i,"cluster"] <- whichCenter
  }
  result
}
```

# Basic Function 3 - Calculate Mean Values of Objects and Update Centers
In this function, the centers will be updated to the mean of clustered objects.
```{r updateCenters}
update <- function(objectsData, centersData){
  #INPUT  objectsData: the data frame with original data and corresponding cluster information
  #INPUT  centersData: all last round data for all centers
  #OUTPUT a data frame with new centers
  #FOR EACH CENTER
  result <- centersData
  for (i in 1:nrow(centersData)){
    #GET ALL NODES IN THIS CLUSTER
    clusterData <- subset(objectsData, objectsData[,"cluster"] == i)
    #CALCULATE MEAN FOR EACH FEATURE & UPDATE CENTERS
    for (j in 1:ncol(centersData)){
      result[i,j] <- mean(clusterData[,j])
    }
  }
  result
}
```

#First Round
Run the algorithm for the first time.
```{r firstRound}
data1 <- assign(data, centers)
data1
centers1 <- update(data1, centers)
centers1
```
We can find that some data records are devided to belong to cluster 1 and others are belong to cluster 2. Update the centers.

#Second Round
Run the alogrithm for the second time.
```{r secondRound}
data2 <- assign(data, centers)
data2
centers2 <- update(data2, centers)
centers2
```
We can find that the centers are not changed. Thus all objects are divided into two clusters and the final clustering result is already got. The clustering result is:

```{r result}
data2
```

```{r verification, include=FALSE}
library(datasets)
verify <- kmeans(data,center=data[c(1,2),])
verify
```