---
title: "kNN Simulation"
author: "MA Mingyu 14110562D"
date: "4/4/2017"
output: pdf_document
---
#COMP4433 Assignment 2 Question 2 a, b and c
##derek.ma@connect.polyu.hk [derek.ma](http://derek.ma)

# Import Training Data
```{r setup}
library(dplyr)
survey <- read.csv("~/Google Drive/_DM/2_Assignments/Ass2/data_q2.csv", stringsAsFactors = FALSE)
colnames(survey) <- c ("customerNo",
                  "averageMonthlyPayment",
                  "averageDurationOfCalls",
                  "totalCallingTime","decision")
```

# Basic Function - Calculate Distance
```{r distance}
distance <- function(vector1, vector2){
  #Euclidean distance
  count <- 0
  for (i in 1:length(vector1)){
    count = count + (vector1[i] - vector2[i])^2
  }
  count^(1/2)
}
```

# Set Value of k and Testing Data
```{r setTestData}
k <- 5
testData <- c(293.26,16.96,120.25)
```

# Preprocessing the Data, Normalization
```{r preprocessing}
#Normalization
min1 <- min(survey$averageMonthlyPayment)
max1 <- max(survey$averageMonthlyPayment)
min2 <- min(survey$averageDurationOfCalls)
max2 <- max(survey$averageDurationOfCalls)
min3 <- min(survey$totalCallingTime)
max3 <- max(survey$totalCallingTime)
testDataNormalized <- c(
  (testData[1]-min1)/(max1-min1),
  (testData[2]-min2)/(max2-min2),
  (testData[3]-min3)/(max3-min3))
survey <- survey %>%
  mutate(averageMonthlyPaymentNormalized = (averageMonthlyPayment - min1)/(max1-min1)) %>%
  mutate(averageDurationOfCallsNormalized = (averageDurationOfCalls - min2)/(max2-min2)) %>%
  mutate(totalCallingTimeNormalized = (totalCallingTime - min3)/(max3-min3))
survey
```

# Calcualte Distance and Sort
```{r calculate distance}
survey <- survey %>%
  mutate(dist=NA)

for (i in 1:length(survey$customerNo)){
  survey[i,"dist"] <- distance(testDataNormalized, 
                      c(survey[i,"averageMonthlyPaymentNormalized"],
                      survey[i,"averageDurationOfCallsNormalized"],
                      survey[i,"totalCallingTimeNormalized"]))
}

surveySorted <- survey[order(survey$dist),]
surveySorted[1:k,]
sort(table(surveySorted[1:k,]$decision))
```

# Calcualte Distance and Sort for Another Test Data
Without considering "Decision", we try quesiont 2b.
```{r calculate distance 2}
testData <- c(271.48,184)
survey <- survey %>%
  mutate(dist=NA)
testDataNormalized <- c(
  (testData[1]-min1)/(max1-min1),
  (testData[2]-min3)/(max3-min3))

for (i in 1:length(survey$customerNo)){
  survey[i,"dist"] <- distance(testDataNormalized, 
                      c(survey[i,"averageMonthlyPaymentNormalized"],
                      survey[i,"totalCallingTimeNormalized"]))
}

surveySorted <- survey[order(survey$dist),]
surveySorted[1:k,]
sort(table(surveySorted[1:k,]$decision))
```

# Cross Validation and Choose the Value of k
```{r cross validation}
for (testIndex in 1:nrow(survey)){
  testData <- c(survey[testIndex,"averageMonthlyPaymentNormalized"],
                survey[testIndex,"averageDurationOfCallsNormalized"],
                survey[testIndex,"totalCallingTimeNormalized"])
  survey <- survey %>%
    mutate(dist=NA)
  for (i in 1:length(survey$customerNo)){
    if(i != testIndex){
      survey[i,"dist"] <- distance(testData, 
                          c(survey[i,"averageMonthlyPaymentNormalized"],
                          survey[i,"averageDurationOfCallsNormalized"],
                          survey[i,"totalCallingTimeNormalized"]))
      surveySorted <- survey[order(survey$dist),]
      #print(tail(names(sort(table(surveySorted[1:k,]$decision))), 1))
      print(sort(table(surveySorted[1:k,]$decision)))
    }
  }
}
```

